Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-02-09 paper/structureScene.tex
new results

\subsection{Gammatone wavelets}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{gfx/gammatones}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
Oscillations (red, blue) are the real and imaginary parts. The envelope (yellow) is the complex modulus.}
\end{center}
\end{figure}
Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment \cite{Venkitaraman2014}, and are shown in Figure \ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelopes of Gammatone wavelets are more biologically plausible than the symmetric, Gaussian envelopes of the more widely used Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.
The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
Likewise, because critical bands are skewed towards higher frequencies, a masker tone has to be louder if it is above the stimulus in frequency rather than below.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{gfx/compression}
\caption{
\label{fig:histograms}
Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
(a) before and (b) after logarithmic compression.}
\label{fig:compression}
\end{center}
\end{figure}

@@

The metric $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items. Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item. The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene. The parameters of the BOF approach are the one used in \cite{aucouturier2007bag}, \texit{i.e.} 10 Gaussians per scene estimated using the Expectation Maximization algorithm. Gaussian models are compared using Monte-Carlo sampling approaximation.

@@

@@

As can be seen on Fiugre \ref{fig:ASS_0}, the logarithmic compression is beneficial for precision performance irrespective of the similarity considered. The gain is more pronounced when considering alternative similarities.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/unsupervised_test1}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$. \ml{ob a missing ??}}
\label{fig:ASS_0}
\end{center}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/unsupervised_test2}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$. \ml{add gmms}}
\label{fig:ASS_1}
\end{center}
\end{figure}

@@

% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lostanlen}}]{Vincent Lostanlen} was born in 1992. He received an engineering degree from TELECOM ParisTech in 2013 and his M.S. degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre et Marie Curie (UPMC) and the Ircam in Paris, France, in 2013. Since 2013, he is a Ph.D. student at \'{E}cole normale sup\'{e}rieure in Paris, working under the supervision of St\'{e}phane Mallat. His research focuses on defining convolutional operators in the time-frequency domain with applications in audio classification.
% \end{IEEEbiography}
%
% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in 1990. He received the B.S. degree in Acoustic from the University Pierre and Marie Curie (UPMC), Paris, France, and the B.S. degree in Musicology from the Sorbonne University, Paris, France, in 2011. He received his M.S. degree in acoustics, signal processing, and musical informatics (ATIAM) from the UPMC and the Ircam in Paris, France, in 2013. Since 2013, he is a Ph.D. student at IRCCyN, Nantes, France. His research interests include acoustic scene similarity and classification as well as acoustic scene perception.
% \end{IEEEbiography}
%
%
% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/anden}}]{Joakim And\'en}
% received the M.Sc. degree in mathematics from the Universit\'e Pierre et Marie Curie, Paris, France, in 2010 and the Ph.D. degree in applied mathematics from Ecole Polytechnique, Palaiseau, France, in 2014. He is currently a postdoctoral researcher with the Program in Applied and Computational mathematics at Princeton University, Princeton, USA. His research interests include signal processing, machine learning, and statistical data analysis. He is a member of the IEEE.
%
% \end{IEEEbiography}
%
% \begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics. He obtained his Ph.D. in computer science at the University of Bordeaux in 2004, and visited several institutions in Canada (University of Victoria, McGill University) and in France (Orange Labs, TELECOM ParisTech, Ircam). His research focuses on machine listening algorithms applied to the analysis of musical and environmental audio.
% \end{IEEEbiography}

>>>>

\subsection{Gammatone wavelets}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/gammatones}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
Oscillations (red, blue) are the real and imaginary parts. The envelope (yellow) is the complex modulus.}
\end{center}
\end{figure}
Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment \cite{Venkitaraman2014}, and are shown in Figure \ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelopes of Gammatone wavelets are more biologically plausible than the symmetric, Gaussian envelopes of the more widely used Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.
The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
Likewise, because critical bands are skewed towards higher frequencies, a masker tone has to be louder if it is above the stimulus in frequency rather than below.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/compression}
\caption{
\label{fig:histograms}
Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
(a) before and (b) after logarithmic compression.}
\label{fig:compression}
\end{center}
\end{figure}

@@

The metric $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items. Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item. The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene. The parameters of the BOF approach are the one used in \cite{aucouturier2007bag}, \textit{i.e.} 10 Gaussians per scene estimated using the Expectation Maximization algorithm. Gaussian models are compared using Monte-Carlo sampling approaximation.

@@

\subsubsection*{Baselines}

As can be seen on Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when considering MFCCs as features. Its performance decreases compared to the early approach when considering the scattering features. Extensive analysis of those results is out of the scope of this paper, but we can assume that this decrease is to to the dimensionality reduction operated on the scattering features before GMM training in order to ensure the convergence of the EM algorithm.

The early approach being simpler in terms of implementation and runtime complexity, this method is retained as baseline for the remaining.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/baselines}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
\label{fig:ASS_-1}
\end{center}
\end{figure}

@@

As can be seen on Figure \ref{fig:ASS_0}, the logarithmic compression is beneficial for precision performance irrespective of the similarity considered. The gain is more pronounced when considering alternative similarities.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/log}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
\label{fig:ASS_0}
\end{center}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/overall}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$. \ml{add gmms}}
\label{fig:ASS_1}
\end{center}
\end{figure}

@@

% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/lostanlen}}]{Vincent Lostanlen} was born in 1992. He received an engineering degree from TELECOM ParisTech in 2013 and his M.S. degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre et Marie Curie (UPMC) and the Ircam in Paris, France, in 2013. Since 2013, he is a Ph.D. student at \'{E}cole normale sup\'{e}rieure in Paris, working under the supervision of St\'{e}phane Mallat. His research focuses on defining convolutional operators in the time-frequency domain with applications in audio classification.
% \end{IEEEbiography}
%
% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/lafay}}]{Gr\'egoire Lafay} was born in 1990. He received the B.S. degree in Acoustic from the University Pierre and Marie Curie (UPMC), Paris, France, and the B.S. degree in Musicology from the Sorbonne University, Paris, France, in 2011. He received his M.S. degree in acoustics, signal processing, and musical informatics (ATIAM) from the UPMC and the Ircam in Paris, France, in 2013. Since 2013, he is a Ph.D. student at IRCCyN, Nantes, France. His research interests include acoustic scene similarity and classification as well as acoustic scene perception.
% \end{IEEEbiography}
%
%
% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/anden}}]{Joakim And\'en}
% received the M.Sc. degree in mathematics from the Universit\'e Pierre et Marie Curie, Paris, France, in 2010 and the Ph.D. degree in applied mathematics from Ecole Polytechnique, Palaiseau, France, in 2014. He is currently a postdoctoral researcher with the Program in Applied and Computational mathematics at Princeton University, Princeton, USA. His research interests include signal processing, machine learning, and statistical data analysis. He is a member of the IEEE.
%
% \end{IEEEbiography}
%
% \begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{figures/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics. He obtained his Ph.D. in computer science at the University of Bordeaux in 2004, and visited several institutions in Canada (University of Victoria, McGill University) and in France (Orange Labs, TELECOM ParisTech, Ircam). His research focuses on machine listening algorithms applied to the analysis of musical and environmental audio.
% \end{IEEEbiography}
